---
title: "Lab 8 - ANOVA, chi-square, beeswarm"
author: "Amelia Ritger"
date: "11/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(car)
library(janitor)
library(ggbeeswarm)
library(kableExtra)
```

#### 1. read in data (penguins.csv)
```{r}
peng <- read_csv("penguins.csv") %>%
  janitor::clean_names()
```

#### 2. find counts of penguins by sex and species
```{r}
peng_counts <- peng %>%
  count(species,sex)
peng_counts
```

#### 3. compare flipper sizes for female penguins across the 3 species
```{r}
peng_f <- peng %>%
  filter(sex=="FEMALE")

#faceted histogram
ggplot(data=peng_f, aes(x=flipper_length_mm)) +
  geom_histogram() +
  facet_wrap(~species)

#faceted qq plot
ggplot(data=peng_f, aes(sample= flipper_length_mm)) +
  geom_qq() +
  facet_wrap(~species)
```

### 4. Find summary statistics
```{r}
peng_f_summary <- peng_f %>%
  group_by(species) %>%
  summarize(mean_flipper_length = mean(flipper_length_mm),
            sd_flipper_length = sd(flipper_length_mm),
            sample_size = n(),
            se_flipper_length = sd(flipper_length_mm)/sqrt(n()),
            varflipper_length = var(flipper_length_mm))
```


Levene's Test (hypothesis test for equal variances):
```{r}
# H0: group variances are equal
leveneTest(flipper_length_mm~species, data=peng_f)
```

Conclusion: Retain the null hypothesis that variances are equal.

Ok, I've done my exploratory analysis, and ALL of these pieces combined are making me think that a one-way ANOVA is an appropriate way to compare means between the 3 species.

#### 5. Make a beeswarm plot with the mean and standard deviation overlayed on top of it

The three most common error bars:

- standard deviation (measure of spread)

- standard error (how far our estimate is from true population)

- confidence interval 

```{r}
ggplot() +
  geom_beeswarm(data=peng_f,
                aes(x=species, y=flipper_length_mm),
                size=1,
                alpha=0.6,
                color="gray60") +
  scale_x_discrete(labels=c("Adelie", "Chinstrap", "Gentoo")) +
  geom_point(data=peng_f_summary,
             aes(x=species, y=mean_flipper_length),
             color="red",
             size=2) +
  geom_errorbar(data=peng_f_summary,
                aes(x=species,
                    ymin=mean_flipper_length - sd_flipper_length,
                    ymax=mean_flipper_length + sd_flipper_length), 
                width=0.1)
```

** NOTE** Make sure that if you're showing summary statistics

**Fig 1** Flipper length (mm) for three female penguin species observed at Palmer Station LTER, Antartica. Gray points indicate individual flipper lengths, red dots indicate group means, error bars indicate $\pm$ 1 standard deviation. Data source:

#### One-way ANOVA

Is there a significant difference in mean flipper length for female penguins between the three species?
```{r}
peng_aov <- aov(flipper_length_mm ~ species, data=peng_f)
summary(peng_aov)
```

H0: Means are equal across all groups
HA: Means are not equal across all groups (at least two differ)

Conclusion: The mean flipper length for the three species for female penguins are NOT equal.

... so then which ones differ significantly?

Answer that with a post-hoc pairwise comparisons test!
```{r}
TukeyHSD(peng_aov)
# diff is the actual difference in means between the two groups
# lwr and upr represent 95% CI on difference between groups
# p adj tells you if there is a sig diff between each pair being compared
```

One-way ANOVA with post-hoc Tukey's HSD reveals significant differences in mean flipper length across all species (Gentoo, mean =..., Adelie, mean=..., ...) (F(df)=..., p <0.001 for al pairwise comparisons).

**Comparing this with the beeswarm plot above... this is good evidence for with large sample sizes, you can find significant ANYTHING. Thus it's important to also include effect sizes, etc!**

#### 7. Making nicer tables (with proportions)
```{r}
party_relig <- gss_cat %>%
  filter(partyid %in% c("Strong republican", "Strong democrat")) %>%
  filter(relig %in% c("None", "Catholic")) %>%
  count(partyid, relig)

# check to see what observations you have retained after filtering, etc
unique(party_relig$partyid)

party_relig_table <- party_relig %>%
  pivot_wider(names_from=relig, values_from=n)
  
```

#### 8. Make a nice table of proportions using janitor::adorn_*
```{r}
party_relig_props <- party_relig_table %>%
  adorn_percentages(denominator="row") %>% #convert to percentages
  adorn_pct_formatting(digits=1) %>% #dictate sig figs
  adorn_ns(position="front") #show counts follow by percentages

party_relig_props #boo! it doesn't look as good as I wanted it to!

kable(party_relig_props) %>%
  kable_styling()
```

Here are a few different ways to ask a question for this scenario:

- Are political and religious affiliation independent from each other?

- Is there a significan association between political affiliation and religious affiliation?

- Is there a significant effect of political association (strong republican vs strong democrat) on religoius affiliation (none, catholic)?

- Is there a significant difference in religoius affiliation between strong republicans and strong democrats?

**AKA are the proportions that exist across groups so different that we think the two groups are not independent of one another?**

```{r}
# Make this a contingency table that ONLY contains the actual counts
chi_counts <- party_relig_table %>%
  select(-partyid)
```

H0:

- There is no signfiicant association between party affiliation and religious affiliation

- There is no significant difference in religious affiliation between the two party groups (str repub or str dem)

- Party and religious affiliation are independent

```{r}
my_party_chi <- chisq.test(chi_counts)
```

Conclusion:

- There is a significant difference in religious affiliation (none, catholic) between strong republicans and strong democrats ($\chi^2$=..., p<0.001)...

- There is a significant association between party ID and religious affiliation

- Party affiliation and religious affiliation are dependent!

